---
title: "Tutorial: Using globaltoolbox to standardize locations and incorportate global data"
author: "Shaun Truelove and Joshua Kaminski"
date: "August 12, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The globaltoolbox R package
    
The `globaltoolbox` R package was developed to help with large-scale analysis projects that include numerous, often unstandarized location names. This package is particularly targeted for use with standardizing location names at various administrative levels from unstandardized sources like free-form surveys or large-scale data aggregation projects that involve multiple sources.

This tutorial is meant to demonstrate some of the applications and workflows that `globaltoolbox` for which is well-suited.

To get started, let's install the package and some others needed for other data manipulation.
```{r install_packages}
install.packages("globaltoolbox")
install.packages("dplyr")
install.packages("readr")
install.packages("tidyr")

```

A demonstration dataset called `globaltbx_demo` is included in the package for demonstration and testing.

```{r demo_data}
data(globaltbx_demo)

head(globaltbx_demo)

```



```{r}

country_aliases.csv <- system.file(
  "extdata",
  "country_aliases.csv",
  package = "globaltoolbox"
)
country_aliases <- utils::read.csv(country_aliases.csv, stringsAsFactors = FALSE)
countries <- country_aliases[,1]
gadm_data <- load_gadm(countries = countries, max_depth = Inf, dbname = default_database_filename())

```





