---
title: "Tutorial: Using globaltoolbox to standardize locations and incorportate global data"
author: "Shaun Truelove and Joshua Kaminsky"
date: "August 12, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# globaltoolbox: An R package for standardizing spatial locations and data
    
The `globaltoolbox` R package was developed to help with large-scale analysis projects that include numerous, often unstandarized location names. This package is particularly targeted for use with standardizing location names at various administrative levels from unstandardized sources like free-form surveys or large-scale data aggregation projects that involve multiple sources.

This tutorial is meant to demonstrate some of the applications and workflows that `globaltoolbox` for which is well-suited.

## Installation

Install the globaltoolbox package from github:

```{r}
devtools::install_github("HopkinsIDD/globaltoolbox")
```

## Use

There are three main types of use of the globaltoolbox package. These include:

1. General country-level location standardization, with fuzzy matching, standardized code identification (i.e., ISO, UN, etc.), and metadata linking and loading.
1. Location standardization at multiple administrative levels using the predefined databases from GADM, GAUL, and others, along with metadata linking, etc.
1. Location standardization and heirarchicalization of user-provided dataset.


For General 

### Create

Before loading the database, please create it with the `create_database` function.

### Load

globaltoolbox provides a `load_gadm` function, which takes the ISO 3166-1 alpha 3 code of a country.  This will download all locations from GADM and add them to the database.

```{r}
globaltoolbox::load_gadm(countries = c('MWI','TZA'))
```

### Search

globaltoolbox provides a `telescoping_standardize` function, which takes your best guess at the name of a location, and returns the standardized name (if it can find it). 

```{r}
# Baltimore, with know country
globaltoolbox::telescoping_standardize(c("unitedstates::baltimore"))
# Just Baltimore, no other information
globaltoolbox::telescoping_standardize(c("baltimore"))
# Misspelled
globaltoolbox::telescoping_standardize(c("baltmore"))


globaltoolbox::telescoping_standardize(c("afganistan::Kuranwamunjan"))
globaltoolbox::telescoping_standardize(c("AFG::Kuranwamunjan"))
globaltoolbox::telescoping_standardize(c("AFG::Kurawamujan"))
globaltoolbox::telescoping_standardize(c("Kuranwamunjan"))

globaltoolbox::standardize_name(c("Kuranwamunjan"))



globaltoolbox::telescoping_standardize("togo::kara")
globaltoolbox::telescoping_standardize("togo")

globaltoolbox::standardize_name(c("togo"))


globaltoolbox::standardize_name(c("togo", "baltimore"))
globaltoolbox::telescoping_standardize(c("togo", "baltimore"))



location_name <- c("usa::baltimore", "togo")

# try multiple
globaltoolbox::telescoping_standardize(c("togo", "baltimore", "kansas", "paris", "Cairo"))

# try multiple, with some scope info
globaltoolbox::telescoping_standardize(c("togo", "unitedstates::baltimore", "unitedstates::kansas", "france::paris", "Cairo"))

# try multiple, with some scope info, and misspellings
globaltoolbox::telescoping_standardize(c("togo", "baltmore", "unitedstates::kanss", "france::paris", "egytp::Kairo"))
# FAILS
globaltoolbox::standardize_name(c("togo", "baltmore", "unitedstates::kanss", "france::paris", "egytp::Kairo"))

location_name <- c("togo", "baltmore", "unitedstates::kanss", "france::paris", "egytp::Kairo")


```



A demonstration dataset called `testingdata_denguetycho` is included in the package for demonstration and testing.

```{r demo_data}
data(testingdata_bgdairportsurvey)
head(testingdata_bgdairportsurvey)


globaltoolbox::telescoping_standardize(c(testingdata_bgdairportsurvey$airports))


```



## Implementation

Standardization of location names/identities is done through integration of multiple global spatial databases (GADM, GAUL), in addition to other datasets of cities and aliases. 
Globaltoolbox maintains a database representing locations including:
locations (numeric id, standard name, readable name, metadata)
Locations represent a named place that is somewhat consistent through time
They are the most important part of the database, and everything else references one or more of them
They are colon separated to show hierarchical relationships
e.g., ::tanzania
aliases (location id, alias)
Aliases are alternate names for locations
They are used primarily to help fix non-standard names
e.g., tza is an alias for ::tanzania
Include alternative standardized codes (ISO, UN, etc.)
geometries (location id, time left, time right, geometry)
Geometries are polygon or point representations of the location
Since the polygon may change in time due to border changes, we store each polygon with a time range representing when it is valid
e.g., https://gadm.org/maps/TZA.html is a geometry for ::tanzania
hierarchy (ancestor location id, last location id, depth)
Hierarchies represent containment relationships
We use hierarchies internally to determine scope (all locations contained in a particular location), and to help correct names
e.g., ::tanzania::pembanorth is contained in ::tanzania

## Main User Functionality

Data Entry Functions
database_add_location_alias
database_add_descendent
load_hierarchical_sf
load_gadm
Data Lookup Functions
get_location_metadata
get_all_aliases
get_location_geometry
(more coming soon)
Name standardization functions
standardize_name
telescoping_standardize

## Name Matching Tech
Name standardization
Encoding and transliteration (


## String Matching Algorithms
Fuzzy matching (multiple used in conjunction)
Match probability -- to be implemented, probability given location level and available locations within database
Typos : telescoping_standardize("tanzani") -> “::tanzania”
Misspelling : telescoping_standardize("tanzinia") -> “::tanzania”
Nonstandard : telescoping_standardize("tza::kusiniunguja") -> "::tanzania::zabzibarsouthandcentral" 
depth first search
> telescoping_standardize("tanzania::chake") -> “::tanzania::pembasouth::chake"

## Additional functionality
Functions for adding other publically available metadata, for example:
WHO incidence data (measles)
Others?

## Still in Progress
Name matching not perfect (and probably never will be)
“::spane” -> “sapone” (“::burkinafaso::bazega::sapone”) without depth argument
Should identify “Spain” ideally. Current R fuzzy matching agorithms are based on string length, number of character matches, etc. Phonetic matching relies on consonants. 
Similarly “Usa” -> “japan::oita::usa”
Database not perfect
“::tanzania” and “::unitedrepublicoftanzania” are different locations
“Usa” is not an alias for “unitedstatesofamerica”
Database large (34 GB for database with shapefiles up to 2nd subcountry level; only 34Mb without shapefiles)
Location merging (to solve the two tanzanias problem). The function exists, but it’s not passing tests yet.
Performance issues (on the full database, matching takes 5 seconds or so)
If possible, use user input to collect additional aliases/alternate names/spellings


## Ideas from the group

Is there specific functionality that others would find valuable?
What use do people have for this sort of package?
Does anyone have good testing data for this? This can be:
Testing data: data with many, non-standardized location names
Training data: data with non-standardized location names AND the fixed/standardized version
Bug reports: a time where you tried to look up something, but got NA or something else instead (include what you looked up, what you got, and what you expected)
Does anyone have data they think would be cool for us to incorporate into the database (lists of alternative names for places, lists of new places, lists of shapefiles for places etc)?
 Integration with other data?
DHS, worldpop, others?






